Investigating social change during cultural contact period using geometric morphometry of pottery shapes from Iron Age northeastern Taiwan

### 2020-02-12 analysis/paper/paper.Rmd
add sentence to methods and test relationship between mean and sd

By standardizing standard deviation in each data set, the coefficients of variation (CV) values allows us to directly compare variation in samples measured with different units or means. This is useful to compare the degree of standardization for archaeological assembles since CV scales standard deviation to the mean that enables comparison of variation across different sample sizes [@Eerkens2001, pp.498]. We take this as our measurement of standardization in vessel shape variables: lower CV values reflect higher standardization, and thus increased craft specialization in the community.

@@

```{r PCs-distribution, fig.cap= "The distribution of PC scores by phases"}
# what are the variences for each PC in each period?
# CVs on shape PCs for standardization
len <- length(op.p$x[,1])
# plot PC1-3 distribution by phase
pc_plot <- tibble(pc = c(op.p$x[,1],
                         op.p$x[,2],
                         op.p$x[,3]),
                  phase = rep(op.p$fac$phase, 3),
                  pcn = str_glue('PC{sort(rep(1:3, len))}'))

@@

```{r CV-test-PCs-table}
pc1 <- data.frame(
x = op.p$x[,1],
y = as.character(op.p$fac$phase),
stringsAsFactors = F)

>>>>

By standardizing standard deviation in each data set, the coefficients of variation (CV) values allows us to directly compare variation in samples measured with different units or means. This is useful to compare the degree of standardization for archaeological assembles since CV scales standard deviation to the mean that enables comparison of variation across different sample sizes [@Eerkens2001, pp.498]. We take this as our measurement of standardization in vessel shape variables: lower CV values reflect higher standardization, and thus increased craft specialization in the community. Given that CV is robust for positive values due to the representation by ratio, we normalized PC scores to a range between 0 and 1 before computation of CV.

@@

```{r PCs-distribution, fig.cap= "The distribution of PC scores by phases"}

# function for normalized data
normalize <- function(x) (x-min(x))/(max(x)-min(x))

# for one matrix
op.p_tmp <- 
op.p$x %>% 
  as_tibble() %>% 
  mutate_all(normalize) %>% 
  as.matrix()

row.names(op.p_tmp) <- row.names(op.p$x )

op.p$x <- op.p_tmp

# what are the variences for each PC in each period?
len <- length(op.p$x[,1])
# plot PC1-3 distribution by phase
pc_plot <- tibble(pc = c(op.p$x[,1],
                         op.p$x[,2],
                         op.p$x[,3]),
                  phase = rep(op.p$fac$phase, 3),
                  pcn = str_glue('PC{sort(rep(1:3, len))}'))

@@

```{r CV-test-PCs-table}
# CVs on shape PCs for standardization
pc1 <- data.frame(
x = op.p$x[,1],
y = as.character(op.p$fac$phase),
stringsAsFactors = F)

### 2020-02-12 analysis/paper/scraps.Rmd
proof from introduction to method section
add sentence to methods and test relationship between mean and sd

>>>>

CV is more robust in the condition that the mean-standard deviation relationship is linear and positive [@Eerkens2001, pp.498]

```{r mean-sd-relationship, eval = FALSE}
# mean and sd by phases for PC scores
pc1_mean_sd_by_phase <-
  pc1 %>% 
  group_by(y) %>% 
  summarize(mean = mean(x),
            sd = sd(x))

pc1_mean_sd_plot <-
  pc1_mean_sd_by_phase %>% 
  ggplot(aes(mean, sd)) +
  geom_point()

pc2_mean_sd_by_phase <-
  pc2 %>% 
  group_by(y) %>% 
  summarize(mean = mean(x),
            sd = sd(x))

pc2_mean_sd_plot <-
  pc2_mean_sd_by_phase %>% 
  ggplot(aes(mean, sd)) +
  geom_point()

pc3_mean_sd_by_phase <-
  pc3 %>% 
  group_by(y) %>% 
  summarize(mean = mean(x),
            sd = sd(x))

pc3_mean_sd_plot <-
  pc3_mean_sd_by_phase %>% 
  ggplot(aes(mean, sd)) +
  geom_point()
```
